\documentclass[12pt]{amsart}

\usepackage{epsf}
\usepackage{geometry}
\usepackage{listings} 
\usepackage{algorithmic,algorithm}
\usepackage[notcite,notref]{showkeys}
\usepackage{multirow}
\usepackage{enumerate}

\usepackage[pdftex]{graphicx}
\usepackage{amscd}
\usepackage[pdftex]{color} % black,white,red,green,blue,cyan,magen ta,yellow
\usepackage[pdftex,colorlinks]{hyperref}
\usepackage{graphicx,pst-eps,epstopdf}
\usepackage{lscape}
\usepackage{indentfirst}
\usepackage{latexsym}
\usepackage{amsmath, amsfonts, amssymb,mathrsfs}
\usepackage{subfigure,pstricks,pst-node}
\usepackage{pst-eps,epstopdf}
\usepackage{verbatim}
%\usepackage{showkeys}

\newenvironment{plan}
{\bigskip\hrule\bigskip\centerline{\bf PLAN}\begin{quote}\tt}
{\end{quote}\bigskip\hrule\bigskip}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in Acrobats bookmarks
    pdftoolbar=true,        % show Acrobats toolbar?
    pdfmenubar=true,        % show Acrobats menu?
    pdffitwindow=true,      % page fit to window when opened
    pdftitle={},    % title
    pdfauthor={Xiaozhe Hu},     % author
    pdfsubject={},   % subject of the document
    pdfnewwindow=true,      % links in new window
    pdfkeywords={}, % list of keywords
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}

\geometry{letterpaper, margin=1in}
\linespread{1.1}

\numberwithin{equation}{section} 
\numberwithin{table}{section}
\numberwithin{figure}{section}
\numberwithin{algorithm}{section}

\everymath{\displaystyle}

\begin{document}

\title[]{Math 226, Homework 3, Due Nov. 13, 2015}
%\author{}
%\date{}

\maketitle


For problems that include programming, please include the code and all outputted figures and tables.  Please label these clearly and refer to them appropriately in your answers to the questions.

\


\begin{enumerate}

\item Show that Jacobi method converges for all $2$ by $2$ symmetric positive definite matrices. 

\

\item \begin{enumerate}
	\item Implement (all the programs should be functions that take matrix $A$, right hand side $b$, initial guess $x^0$, tolerance and maximal number of iterations as input arguments and output the approximate solution and number of iterations): 
		\begin{enumerate}
		\item Richardson method
		\item Jacobi method
		\item Guass-Seidel method
		\item Successive Overrelaxation (SOR) method
		\end{enumerate}
 	\item Consider the following $n \times n$ system:
		$$\left(%
	\begin{array}{ccccc}
	  2 & -1 &  &  &  \\
	  -1 & 2 & -1 &  &  \\
	   & \ddots & \ddots & \ddots &  \\
	   &  & -1 & 2 & -1 \\
	   &  &  & -1 & 2 \\
	\end{array}%
	\right)
	\left(%
	\begin{array}{c}	
	  x_1 \\
	   \\
	   \vdots\\
	   \\
	  x_n \\
	\end{array}%
	\right)=
	\left(%
	\begin{array}{r}
	  1 \\
	  0 \\
	  \vdots \\
	  0 \\
	  -1 \\
	\end{array}%
	\right), 
	$$	
	for  $n=2^l$, $l=4,5,6,7,8$. Use the programs above to solve it until $|| b - A x^k ||_2/ ||b||_2 < 10^{-6}$. Make a table to report the number of iterations of each iterative method and discuss the results.  For Richardson method and SOR, please try different values of $\omega$ and discuss how the results depends on the choice of $\omega$.
	\end{enumerate}
	
\

\item Verify the Shern-Morrison-Woodbury formula. If $\widetilde{\mathbf{A}} = \mathbf{A} + \mathbf{uw}^{T}$, then
$$
\widetilde{\mathbf{A}}^{-1} = \mathbf{A}^{-1} - \frac{1}{1+\mathbf{w}^{T} \mathbf{A}^{-1} \mathbf{u}} \mathbf{A}^{-1} \mathbf{u w}^{T} \mathbf{A}^{-1}
$$


\

\item Consider
\vskip -20pt
\begin{align*}
x_{1}^{2} + x_{2}^{2} &= 1 \\
(x_{1}-1)^{2} + x_{2}^{2} & =1 
\end{align*}
\begin{enumerate}
\item Implement Newton's Method and find all solutions. 
\item Implement Broyden I Method and use $B_{0} = I$ to find all solutions.
\item Implement Broyden II Method and use $H_{0} = I$ to find all solutions.
\end{enumerate}

\end{enumerate}

\end{document}